---
title: "Learning - pilot"
author: "Matan Mazor"
date: "06/07/2020"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
library("papaja")
library("tidyverse")
library("broom")
library("cowplot")
library('MESS') # for AUCs
library('lsr') # for effect sizes
library('pwr') # for power calculations
library('brms') # for mixed effects modeling
```


```{r pilot1, echo=FALSE, cache=TRUE, include=FALSE}
Lotz_FP <- read_csv('../Experiments/Lotz/results/pilot_FP/jatos_results_batch1.csv') %>%
  rename('subj_id' = 'PROLIFIC_PID') %>%
  mutate(subj_id=factor(subj_id),
         num_items = 1)
FP_N <- Lotz_FP$subj_id%>%unique()%>%length()

# PAY BONUS
Lotz_FP %>%
  select(c(subj_id,bonus)) %>%
  group_by(subj_id) %>%
  summarise(bonus=mean(bonus)) %>%
  write.table('../Experiments/Lotz/results/pilot_FP/bonus_batch1.csv', sep=',', row.names=FALSE, col.names=FALSE,quote=F)

Lotz_FN <- read_csv('../Experiments/Lotz/results/pilot_FN/jatos_results_batch1.csv') %>%
  rename('subj_id' = 'PROLIFIC_PID') %>%
  mutate(subj_id=factor(subj_id),
         num_items = 1)

FN_N <- Lotz_FN$subj_id%>%unique()%>%length()
# PAY BONUS
Lotz_FN %>%
  select(c(subj_id,bonus)) %>%
  group_by(subj_id) %>%
  summarise(bonus=mean(bonus)) %>%
  write.table('../Experiments/Lotz/results/pilot_FN/bonus_batch1.csv', sep=',', row.names=FALSE, col.names=FALSE,quote=F)

# Concatenate the two dfs
Lotz_df <- rbind(Lotz_FP,Lotz_FN)%>%
  filter(test_part=='test' & trial_part=='main') %>%
  select(c('trial_index','subj_id','FP','bonus','effect','rt','key_press','correct','trial_name')) %>%
  group_by(subj_id) %>%
  mutate('trial_index' = rank(trial_index),
         'block' = floor((trial_index-1)/8)+1,
         'include' = rt>250,
         'stimulus_pair'= case_when(
           trial_name == 'Q+' | trial_name =='O-'   ~ 'Q_O',
           trial_name == 'Q-' | trial_name =='O+'   ~ 'O_Q',
           trial_name == 'A-' | trial_name =='B+'   ~ 'B_A',
           trial_name == 'C+' | trial_name =='CD-'   ~ 'C_CD',
           trial_name == 'EF+' | trial_name =='E-'   ~ 'EF_E'
         ),
         'include'=ifelse(is.na(include),FALSE, include),
         'include_s'=(sum(include)>50 & trial_name%>%is.na()%>%sum()==0)
  ) %>%
  filter(include_s) %>%
  ungroup()

totalN <- Lotz_df %>%
  group_by(FP )%>%
  summarise(N = subj_id %>%
              unique() %>%
              length())

## Overall accuracy

acc_df <- Lotz_df %>%
  group_by(subj_id) %>%
  summarise(acc=mean(correct,na.rm=TRUE)) 


learning_curves <- Lotz_df %>%
  group_by(block, stimulus_pair) %>%
  summarise(accuracy = mean(correct, na.rm=TRUE),
            se_accuracy = sd(correct, na.rm=TRUE)/sqrt(length(correct)),
            RT = median(rt, na.rm=TRUE),
            se_RT = sd(rt, na.rm=TRUE)/sqrt(length(rt)))


# ggplot(learning_curves%>%filter(stimulus_pair=='C_CD' | stimulus_pair=='EF_E'), 
#        aes(x=block, y=accuracy, color=stimulus_pair))+
#   geom_line()+
#   geom_ribbon(aes(ymin=accuracy-se_accuracy, ymax=accuracy+se_accuracy, fill=stimulus_pair), alpha=0.5)
# 
# ggplot(learning_curves%>%filter(stimulus_pair=='C_CD' | stimulus_pair=='EF_E' | stimulus_pair=='Q_O' | stimulus_pair=='O_Q'), 
#        aes(x=block, y=accuracy, color=stimulus_pair))+
#   geom_line()+
#   geom_ribbon(aes(ymin=accuracy-se_accuracy, ymax=accuracy+se_accuracy, fill=stimulus_pair), alpha=0.5)
# 
# ggplot(learning_curves%>%filter(stimulus_pair=='C_CD' | stimulus_pair=='EF_E' | stimulus_pair=='Q_O' | stimulus_pair=='O_Q'), 
#        aes(x=block, y=RT, color=stimulus_pair))+
#   geom_line()+
#   geom_ribbon(aes(ymin=RT-se_RT, ymax=RT+se_RT, fill=stimulus_pair), alpha=0.5)

accuracy_df <- Lotz_df %>%
  filter(stimulus_pair!='Q_O' & stimulus_pair!='O_Q') %>%
  group_by(subj_id, stimulus_pair, block) %>%
  summarise(accuracy = mean(correct, na.rm=TRUE))

accuracy_anova <- aov(accuracy ~ stimulus_pair+block+subj_id, data=accuracy_df) %>%
  summary()

accuracy_by_pair <- Lotz_df %>%
  group_by(subj_id, stimulus_pair) %>%
  summarise(accuracy = mean(correct, na.rm=TRUE))%>%
  spread(stimulus_pair,accuracy)

abp_t1 <- t.test(accuracy_by_pair$B_A,
                (accuracy_by_pair$EF_E+
                   accuracy_by_pair$C_CD)/2,
                paired=TRUE)

abp_t2 <- t.test(accuracy_by_pair$EF_E,
                accuracy_by_pair$C_CD,
                paired=TRUE)

#   spread(stimulus_pair, accuracy)%>%
#   mutate(norm_C_CD=C_CD/B_A,
#          norm_EF_E=EF_E/B_A,
#          norm_O_Q = O_Q/B_A,
#          norm_Q_O = Q_O/B_A)
#   mutate(FP_norm = abs(Q_O-EF_E)-abs(Q_O-C_CD),
#          FN_norm = abs(O_Q-EF_E)-abs(O_Q-C_CD)) %>%
#   ungroup()%>%
#   summarise(FP_norm = mean(FP_norm, na.rm=TRUE),
#             FN_norm = mean(FN_norm, na.rm=TRUE,))

```

# Methods

## Participants

A total of `r FN_N+ FP_N` participants were recruited from Prolific. Participants were Native English Speakers and had a minimum approval rate of 0.9. They were paid £0.80 for their participation, in addition to a bonus payment of up to £0.64. The experiment took 7 minutes to complete.

## Experimental design

On each trial, participants were presented with one or two black shapes, presented inside a frame in the center of the screen. They then used their keyboard to indicate whether they thought a cat will appear after the shape ('J' key) or not ('F' key). Following their response, the shapes disappeared from the screen. On half of the trials, the frame remained empty for 2 seconds. On the other half, an image of a cat appeared inside the frame. Participants were awarded a point for accurate predictions, and points were later translated to a bonus payment with a conversion rate of £0.01 per point. Their total number of point and trial count remained visible on the screen throughout the experiment. 

After Reading the information sheet and signing the consent form, participants were instructed about the task. In the following practice part, participants learned to associate one shape with the appearance of a cat, and one shape with the absence of a cat. Shapes were randomly selected separately for each participant from the following: circle, hexagon, up-facing triangle, down-facing triangle, 5-armed star, heart, spade, club, diamond, and cross. Once they performed with 100% accuracy for a block of four trials, they moved on to a verbal comprehension test ('Please describe the task in your own words'), and then moved on to the main task.

The main task consisted of 64 trials, separated into two parts of 32 trials each. Trials were presented in 8 blocks, each block comprising 8 different trial types, as follows:

1. Stimulus A, coupled with cat presence.
2. Stimulus B, coupled with cat absence.
3. Stimulus C, coupled with cat presence,
4. Stimuli C and D, coupled with cat absence.
5. Stimuli E and F, coupled with cat presence.
6. Stimulus E, coupled with cat absence.

Stimuli A-F were randomly sampled, without replacement, from the remaining shapes that were not used as example stimuli in the practice phase.

For `r FN_N` of the participants (Q positive group), the two remaining trial-types were:

7. A ring intersected by a line (Q), coupled with cat presence.
8. A ring (O), coupled with cat absence.

For the remaining `r FP_N` participants (Q negative group), the two remaining trial-types were:

7. A ring (O), coupled with cat presence.
8. A ring intersected by a ling (Q), coupled with cat absence.

Trial types 1 and 2 formed the symmetric learning condition, where one stimulus predicted the presence of a cat, and one stimulus predicted its absence. Trial types 3 and 4 formed the Feature-Negative condition, where the presence of a feature (stimulus D) predicted the absence of a cat. Trial types 4 and 5 formed the Feature-Positive condition, where the presence of a feature (stimulus F) predicted the presence of a cat. Finally, trial types 7 and 8 formed the test case, where we tested whether the assignment of stimuli to the presence or absence of a cat makes learning more or less efficient. 

## Exclusion

Participants were excluded for responding too quickly (RT<250) in more than 14 trials. After exclusion, we had datasets from `r totalN[1,2]` participants in the Q positive and `r totalN[2,2]` in the Q negative groups, respectively. 

## Results

Overall accuracy was high (`r apa_print(t.test(acc_df$acc))$estimate`). 
Analysis of variance was applied to the data from trial-types 1-6 (not including the test stimuli), with the dependent variable accuracy and independent variables block number, stimulus pair, and participant. As expected, block number had a significant effect on accuracy, suggesting that participants learned to associate stimuli with outcome over the course of the experiment (`r apa_print(accuracy_anova)$statistic$block`). Furthermore, accuracy was significantly different for the different stimulus pairs (`r apa_print(accuracy_anova)$statistic$stimulus_pair`; see Fig. \ref{fig:learning_curves}).

However, this difference was driven by faster learning for the symmetric learning condition (trial types 1 and 2) compared to the other learning conditions (`r apa_print(abp_t1)$full_result`). No significance difference was observed in overall accuracy between the feature-positive and feature-negative learning accuracies (`r apa_print(abp_t2)$full_result`). 



```{r fig:learning_curves, echo=FALSE, fig.cap='Learning curves for the three within-participant conditions'}
ggplot(learning_curves%>%filter(stimulus_pair=='C_CD' | stimulus_pair=='EF_E' | stimulus_pair=='B_A'),
       aes(x=block, y=accuracy, color=stimulus_pair))+
  geom_line()+
  geom_ribbon(aes(ymin=accuracy-se_accuracy, ymax=accuracy+se_accuracy, fill=stimulus_pair), alpha=0.5)
```

```{r fig:RT_curves, echo=FALSE, fig.cap='Learning curves for the three within-participant conditions'}
ggplot(learning_curves%>%filter(stimulus_pair=='C_CD' | stimulus_pair=='EF_E' | stimulus_pair=='B_A'),
       aes(x=block, y=RT, color=stimulus_pair))+
  geom_line()+
  geom_ribbon(aes(ymin=RT-se_RT, ymax=RT+se_RT, fill=stimulus_pair), alpha=0.5)
```

