---
title: "go/no-go - pilot"
author: "Matan Mazor"
date: "06/07/2020"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
library("papaja")
library("tidyverse")
library("broom")
library("cowplot")
library('MESS') # for AUCs
library('lsr') # for effect sizes
library('pwr') # for power calculations
library('brms') # for mixed effects modeling
library('BayesFactor')
library('afex') # for anova
```


```{r pilot1, echo=FALSE, cache=TRUE, include=FALSE}
gng_df_unfiltered <- read_csv('../Experiments/go-no-go/results/pilot1/jatos_results_batch1.csv') %>%
  rename('subj_id' = 'PROLIFIC_PID') %>%
  mutate(subj_id=factor(subj_id),
         num_items = 1)

Ntotal <- gng_df_unfiltered %>%
  summarise(N = subj_id %>%
              unique() %>%
              length())

# PAY BONUS
gng_df_unfiltered %>%
  select(c(subj_id,bonus)) %>%
  group_by(subj_id) %>%
  summarise(bonus=mean(bonus)) %>%
  write.table('../Experiments/go-no-go/results/pilot1/bonus_batch1.csv', sep=',', row.names=FALSE, col.names=FALSE,quote=F)

# filter
gng_df <- gng_df_unfiltered%>%
  filter(test_part=='test' & trial_part=='main') %>%
  select(c('trial','block','trial_index','subj_id','bonus','rt','key_press','correct','shape_array', 'respond')) %>%
  group_by(subj_id) %>%
  mutate(
    'include_s'=(mean(correct)>0.5),
    'condition' = case_when(
      shape_array=="['stimuli/O.png', 'stimuli/Q.png']" ~ 'FP',
      shape_array=="['stimuli/Q.png', 'stimuli/O.png']" ~ 'FN'
    ),
    'first_block'=ifelse(
      shape_array[1]=="['stimuli/O.png', 'stimuli/Q.png']",'FP','FN'
      ),
    'trial_type'=ifelse(respond==0,'no-go','go')
  ) %>%
  filter(include_s) %>%
  ungroup()

Nincluded <- gng_df %>%
  summarise(N = subj_id %>%
              unique() %>%
              length())


##  accuracy

acc_df <- gng_df %>%
  group_by(subj_id,condition) %>%
  summarise(acc=mean(correct,na.rm=TRUE)) %>%
  spread(condition,acc) %>%
  mutate(acc=(FP+FN)/2)

accxresp_df <- gng_df %>%
  group_by(subj_id,condition,trial_type) %>%
  summarise(acc=mean(correct,na.rm=TRUE)) 

acc_anova <- afex::aov_ez(
  data=accxresp_df,
  dv='acc',
  id='subj_id',
  within=c('trial_type','condition')
  )


RT_df <- gng_df %>%
  group_by(subj_id,condition) %>%
  summarise(RT=mean(rt,na.rm=TRUE)) %>%
  spread(condition,RT)

```

# Methods

## Participants

A total of `r Ntotal$N` participants were recruited from Prolific. Participants were Native English Speakers and had a minimum approval rate of 0.9. They were paid £0.50 for their participation, in addition to a bonus payment of up to £0.32. The experiment took 4 minutes to complete.

## Experimental design

On each trial, participants were presented with a fixation cross (500 ms) followed by one of two shapes (500 ms). They were asked to respond with a key press within 500 ms to one shape ('go' trials), but not to the other ('no-go' trials). Participants were awarded a point for responding on 'go' trials and for not responding on 'no-go' trials. Points were later translated to a bonus payment with a conversion rate of £0.0025 per point. . 

After Reading the information sheet and signing the consent form, participants were instructed about the task. In the following practice part, participants had 8 practice trials in which they responded to a club, but not to a heart. The practice block was followed by a verbal comprehension test ('Please describe the task in your own words'), which was followed by the main task.

The main task consisted of 128 trials, separated into four blocks of 32 trials each. In two blocks, participants responded to the appearance of a ring intersected by a line, but not to the appearance of a ring. In the other two blocks, participants responded to the appearance of a ring, but not to the appearance of a ring intersected by a line. The order of blocks was interleaved. The first block was chosen at random with 50% probability. 

## Exclusion

Participants were excluded for performing below chance. After exclusion, we had datasets from `r Nincluded$N` participants. 

## Results

Overall accuracy was very high (`r apa_print(t.test(acc_df$acc))$estimate`).

Accuracy was higher for blocks in which a response was required for feature presence (FP blocks; `r apa_print(t.test(acc_df$FP))$estimate`) than for feature absence (FN blocks; `r apa_print(t.test(acc_df$FN))$estimate`; `r apa_print(t.test(acc_df$FP,acc_df$FN, paired=TRUE))$full_result`). 

No difference was observed between the reaction times for the two block types (median reaction time in FP blocks: `r printnum(median(RT_df$FP, na.rm=TRUE))`; median reaction time in FN blocks: `r printnum(median(RT_df$FN, na.rm=TRUE))`; t-test on the log-transformed reaction times: `r apa_print(ttestBF(log(RT_df$FN),log(RT_df$FP),paired=TRUE))$full_result`). 

```{r fig_accuracy, echo=FALSE, fig.cap='Accuracy as a function of block and trial'}
ggplot(accxresp_df,aes(fill=trial_type,y=acc,x=condition))+
  geom_boxplot()+
  labs(y='accuracy',x='block type')
```

