---
title: "simulation"
author: "Matan Mazor"
date: "04/08/2020"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(brms)
library(broom)
library(mvtnorm)
```

## R Markdown

For the purpose of illustration only, we assume search asymmetry effect sizes based on @treisman1988feature, @treisman1985search, and  @wang1994familiarity. We quantify the search asymmetry effect size as the difference between the slope for 'no' responses in the feature-absence condition and in the feature-presence condition. In a serial search, this quantity reflects the additional cost in milliseconds per item for searching the feature-absent compared to the feature-present target. Based on our pilot data, our generative model assumes a standard deviation of 48 milliseconds across participates for this measure, for a fixed stimulus pair. 

Next, our model assumes a correlation of -0.25 between search asymmetry and metacognitive asymmetry, measured as the area between the feature-present and feature-absent type-2 response conditional ROC curves. This assumption is not grounded in any observation, but instead marks our smallest effect size of interest (SESOI) for this study. 

```{r echo=FALSE,cache=TRUE}

search_asym = tibble('curvature'=31.8, 
                     'tilt'=27.1, 
                     'color'=3.8, 
                     'gap (C or O)'=31.8, 
                     'part (Q or O)'=34.9, 
                     'letter inversion'=34)

get_search_asym <- function(exp) {
  return(get(exp,search_asym))
}
n <- 200
std <- 48
r <- -0.25

#sample individual data points
# to make sure the correlation is around 0.25, I follow Greg McFarlane's answer from here:
# https://stats.stackexchange.com/questions/15011/generate-a-random-variable-with-a-defined-correlation-to-an-existing-variables
# to restrict ROC asymmetry to be between -1 and 1 I take the scaled cumulative normal distribution of th resulting vector.
d <-
  tibble(
    # Six experiments
    exp = rep(search_asym%>%colnames(), each=n))%>% 
    group_by(exp)%>%
    mutate(
        search_asym = rnorm(n,get(exp,search_asym),std))%>%
    ungroup() %>%
    mutate(
        search_asym_scaled = search_asym/sd(search_asym),
        ROC_asym = (rnorm(n*6,0,1)+search_asym_scaled*1/tan(acos(r))))

summary <- d%>%group_by(exp)%>%summarise(
  ROC_search_corr = cor(search_asym,ROC_asym),
  ROC=mean(ROC_asym),
  search=mean(search_asym))

png(filename='figures/correlations.png', width=7, height=4, units='in', res=300)

ggplot(d,aes(x=ROC_asym,y=search_asym,color=exp,shape=exp)) +
  geom_point(alpha=0.2) + 
  geom_smooth(method = "lm", fill = NA) +
  geom_point(data=d %>%
               group_by(exp) %>%
               summarise_at(vars('ROC_asym','search_asym'),mean),
             size=5)+
  labs(x='metacognitive asymmetry',y='search asymmetry', title='Between-subjects correlation')

dev.off()
```



```{r echo=FALSE,cache=TRUE}
# sample asymmetry means for the four tasks and six stimulus types.
population_means <- mvtnorm::rmvnorm(6,mean=c(0,0,0,0),sigma=diag(4)*0.1+0.9)

n <- 200

#sample individual data points
d <-
  tibble(
    # Six experiments
    exp = rep(c(1,2,3,4,5,6), each=n))%>% 
    group_by(exp) %>% 
     mutate(
       # the beta coefficient relating ROC and search asymmetries
       beta_ROC = rnorm(1,0.3,0.1),
       #ROC asymmetry is bounded to [-1, 1], so I apply pnorm to squeeze it into this interval.
       # if bias>0, ROC_asym is closer to 1. If bias <0, closer to 0.
        ROC_asym = 2*(pnorm(rnorm(n,population_means[exp,1],0.5))-0.5),
        standardized_ROC_asym=(ROC_asym-mean(ROC_asym))/sd(ROC_asym),
       # searc_asym is sampled from a normal distribution centred around bias+beta_ROC*standardized ROC_asym
        search_asym = rnorm(n,population_means[exp,2],1),
       # go-nogo asym
       gng_asym = rnorm(n,population_means[exp,3],1),
       # change detection asym
       cd_asym = rnorm(n,population_means[exp,4],1),
       # the correlation is not used anywhere, and is here only as a sanity check. 
        cor = cor(ROC_asym,search_asym))

summary <- d%>%group_by(exp)%>%summarise(
  ROC_asym=mean(ROC_asym),
  search_asym=mean(search_asym),
  gng_asym=mean(gng_asym),
  cd_asym=mean(cd_asym))

```

```{r load_pilot_data, echo=FALSE, cache=TRUE}

search_df <- read_csv('vs_simulation.csv')%>%
  mutate(target=ifelse(target=='Q',1,2),
         target=factor(target,levels=c(1,2)),
         response=factor(response,levels=c(0,1)),
         RT=search_time)

median_search_times <- search_df %>%
  group_by(set_size,target,response) %>%
  summarise_at(c('correct','RT'),mean) %>%
  mutate(response=ifelse(response==1,'present','absent') %>%
          factor(levels = c('present','absent')), 
         target=ifelse(target==1,'Q-in-O','O-in-Q') %>%
           factor(levels=c('Q-in-O','O-in-Q')))%>%
  group_by(set_size,target,response)%>%
  summarise(median_RT= median(RT), sem_RT=se(RT)*1.2533) 

search_slopes <- search_df %>%
  filter(!subj_id%in%excluded$subj_id & correct==1 & RT>minRT & RT<maxRT) %>%
  group_by(subj_id,target,response) %>%
  do(model=lm(RT~set_size,data=.)) %>%
  tidy(model) %>%
  filter(term=='set_size') %>%
  mutate(response=ifelse(response==1,'present','absent') %>%
          factor(levels = c('present','absent')), 
         target=ifelse(target==1,'Q','O') %>%
           factor(levels=c('Q','O')))

search_slopes_by_target <- search_df %>%
  filter(!subj_id%in%excluded$subj_id & correct==1 & RT>minRT & RT<maxRT) %>%
  group_by(subj_id,target) %>%
  do(model=lm(RT~set_size,data=.)) %>%
  tidy(model) %>%
  filter(term=='set_size') %>%
  mutate(target=ifelse(target==1,'Q','O') %>%
           factor(levels=c('Q','O')))

mean_search_slopes <- search_slopes %>%
  group_by(target,response) %>%
  summarise('mean_slope'=mean(estimate,na.rm=TRUE),
            'se_slope' = se(estimate, na.rm=TRUE))

mean_search_slopes <- search_slopes %>%
  group_by(target,response) %>%
  summarise('mean_slope'=mean(estimate,na.rm=TRUE),
            'se_slope' = se(estimate, na.rm=TRUE))

slopes_by_target <- search_slopes %>%
  group_by(target) %>%     
  summarise('mean_slope'=mean(estimate,na.rm=TRUE)) %>%
  spread(target, mean_slope)

slopes_by_response <- search_slopes %>%
  group_by(response) %>%     
  summarise('mean_slope'=mean(estimate,na.rm=TRUE)) %>%
  spread(response, mean_slope)

anv <- aov(estimate ~ target * response, data = search_slopes)

asymmetry_df <- search_slopes_by_target %>%
  group_by(subj_id,target) %>%
  summarise(slope=mean(estimate)) %>%
  spread(target,slope) %>%
  mutate(asymmetry=(O-Q)/(O+Q))
```


```{r pressure, echo=FALSE}
plot(pressure)
```

